---
title: "Pay Equity Report"
output: pdf_document
params: 
     table: NULL
     graph: NULL
     focal: NULL
     level: NULL
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

This report contains two major sections: a t-test analysis table and a percentage disparity bar chart.

### t-Test Analysis Table

First, a table is displayed that contains the following columns:

1. Pay Group (job title, grade, or however you group the workforce)
1. Number of total employees in the pay group
1. Number of focal group employees in the pay group
1. Average pay of members of the focal group
1. Average disparity - the average difference in pay between the focal group and all other employees
1. t-Test - the statistical test for the difference in pay between focal and other groups
1. p-value - the probability the t-Test value would be observed if the pay process was fair

Note that a p-value < .05 is the generally accepted criterion for a statistically significant difference in pay.  A p-value < .05 means that there is less than a 1 in 20 chance the observed t-test value would occur if the disparity were due to chance (i.e., if the pay process was fair).

### Chart of Disparity as a Percentage

Second, a bar graph is displayed that presents the average disparity as a percent of the other group's pay.  This may be useful because it puts pay groups on extreme ends of the pay spectrum on a common metric.  This enables users to note whether disparity persists across the entire workforce as well.  Bars on the left side of the graph represent disparity against the focal group, while bars on the right side of the graph represent disparity against the other group. 

As you play around with different focal groups, you may note that simply flipping the focal group from one group to another does not simply change the sign of the percentage, it actually changes the value also.  This is because percentages are based on the average pay of the reference (or "other") group.  So for example, if you have a workforce with two genders (men and women), and the analysis shows that women are underpaid by 10%, the application will show that men are overpaid by 11.11% when women are the focal group.

\blandscape

```{r data-table, echo = FALSE}
col_names <- c(params$level, "# EEs", paste("#", params$focal), 
               paste(params$focal, "Avg. Pay"), "Avg. Disparity", "t-Test", "p-value")

disp_tbl <- kable(params$table,
                  caption = "Pay Equity Analysis: t-Test Results by Group",
                  digits = 3,
                  col.names = col_names)
disp_tbl
```

\newpage

```{r result-graph, echo = FALSE}
# for readability, set dynamic x-axis scale limits
scale_max <- params$graph |> 
     summarize(scale_max = max(abs(pct_underpaid))) |> 
     mutate(scale_min = -1 * scale_max)

ggplot(data = params$graph, aes(x = pct_underpaid, y = level, fill = color_group)) +
               ggtitle("Pay Equity Analysis: % Disparity by Group") +
               geom_bar(stat = "identity") +
               geom_text(aes(label = paste(format(pct_underpaid, digits = 3), "%"), 
                             hjust = hjust,
                             fontface = "bold")) +
               ylab(params$level) +
               xlab(paste("%", params$focal, "Underpaid/Overpaid")) +
               scale_x_continuous(limits = c(scale_max$scale_min, scale_max$scale_max)) +
               geom_vline(xintercept = 0) +
               scale_fill_manual(values = c("negative" = "deepskyblue4", 
                                            "positive" = "grey50")) +
               guides(fill = FALSE)
```

\elandscape